RAG1 and RAG2 non-core regions are implicated in the leukemogenesis and off-target V(D)J recombination in BCR-ABL1-driven B cell lineage lymphoblastic leukemia

Version 1

Abstract

Abstract

The evolutionary conservation of non-core RAG regions suggests significant roles that might involve quantitative or qualitative alterations in RAG activity. Off-target V(D)J recombination contributes to lymphomagenesis and is exacerbated by RAG2’ C-terminus absence in Tp53-/- mice thymic lymphomas. However, the genomic stability effects of non-core regions from both cRAG1 and cRAG2 in BCR-ABL1+ B-lymphoblastic leukemia (BCR-ABL1+ B-ALL), the characteristics, and mechanisms of non-core regions in suppressing off-target V(D)J recombination remains unclear. Here, we established three mice models of BCR-ABL1+ B-ALL in full-length RAG (fRAG), core RAG1 (cRAG1), and core RAG2 (cRAG2) mice. The cRAG(cRAG1 and cRAG2) leukemia cells exhibited greatr malignant tumor characteristics compared to fRAG cells. Additionally, cRAG cells showed higher frequency of off-target V(D)J recombination and oncogenic mutations than fRAG. We also revealed decreased RAG binding accuracy in cRAG cells and a smaller recombinant size in cRAG1 cells, which could potentially exacerbate off-target V(D)J recombination in cRAG cells. In conclusion, these findings indicate that the non-core RAG regions, particularly the non-core region of RAG1, play a significant role in preserving V(D)J recombination precision and genomic stability in BCR-ABL1+ B-ALL.

Introduction

V(D)J recombination serves as the central process for early lymphocyte development and generates diversity in antigen receptors. This process involves the double-strand DNA cleavage of gene segments by the V(D)J recombinase, including RAG1 and RAG2. RAG recognizes conserved recombination signal sequences (RSSs) positioned adjacent to V, D, and J gene segments. A bona fide RSS contains a conserved palindromic heptamer (consensus5′-CACAGTG) and A-rich nonamer (consensus 5′ACAAAAACC) separated by a degenerate spacer of either 12 or 23 base pairs (Schatz and Ji, 2011; Hirokawa, et al., 2020). The process of efficient recombination is contingent upon the presence of recombination signal sequences (RSSs) with differing spacer lengths, as dictated by the “12/23 rule” (Eastman, et al., 1996; Banerjee and Schatz, 2014). Following cleavage, the DNA ends are joined via non-homologous end joining (NHEJ), resulting in the precise alignment of the two coding segments and the signaling segment (Rooney, et al., 2004). V(D)J recombination promotes B cell development, but aberrant V(D)J recombination can lead to precursor B-cell malignancies through RAG mediated off-target effects (Thomson, et al., 2020; Mendes, et al., 2014; Onozawa and Aplan, 2012).

The regulation of RAG expression and activity is multifactorial, serving to ensure V(D)J recombination and B cell development (Gan, et al., 2021; Kumari, et al., 2021). The RAGs consist of core and non-core region. Although recombinant dispensability appears evident, the non-core RAG regions are evolutionarily conserved, indicating their potential significance in vivo that may involve quantitative or qualitative modifications in RAG activity and expression (Liu, et al., 2022; Curry and Schlissel, 2008; Sekiguchi, et al., 2001). Specifically, the non-core RAG2 region (amino acids 384–527 of 527 residues) contains a plant homeodomain (PHD) that can recognize histone H3K4 trimethylation, as well as a T490 locus that mediates a cell cycle-regulated protein degradation signal in proliferated pre-B cells stage (Matthews, et al., 2007; Liu, et al., 2007). The off-target V(D)J recombination frequency is significantly higher when RAG2 is C-terminally truncated, thereby establishing a mechanistic connection between the PHD domain, H3K4me3-modified chromatin, and the suppression of off-target V(D)J recombination (Lu, et al., 2015; Mijušković, et al., 2015). RAG2 destruction is linked to the cell cycle through the cyclin-dependent kinase cyclinA/Cdk2, which phosphorylates T490. Failure to degrade RAG during the S stage poses a threat to the genome (Zhang, et al., 2011). A T490A mutation at the phosphorylation site contribute to lymphomagenesis in a p53-deficient background. The RAG1’ non-core region (amino acids1–383 of 1040 residues) has been identified as a RAG1 regulator. While the core RAG1 maintains its catalytic activity, it’s in vivo recombination efficiency and fidelity are reduced in comparison to the full-length RAG1 (fRAG1). In addition, the RAG binding to the genome is more indiscriminate (Silver, et al., 1993; Beilinson, et al., 2021; Sadofsky, et al., 1993). The N-terminal domain (NTD), which is evolutionarily conserved, is predicted to contain multiple zincbinding motifs, including a Really Interesting New Gene (RING) domain (aa 287 to 351) that can ubiquitylate various targets, including RAG1 itself (Deng, et al., 2015). While the ubiquitylation activity has been characterized in vitro, its in vivo relevance to V(D)J recombination and off-target V(D)J recombination remains uncertain. Furthermore, the N-terminal domain (NTD) contains a specific region (amino acids 1 to 215) that facilitates interaction with DCAF1, leading to the degradation of RAG1 in a CRL4dependent manner (Schabla, et al., 2018). Additionally, the NTD plays a role in chromatin binding and the genomic targeting of the RAG complex (Schatz and Swanson, 2011). Despite increased evidence emphasizing the significance of non-core RAG regions, particularly RAG1’s non-core region, the function of non-core RAG regions in off-target V(D)J recombination and the underlying mechanistic basis have not been fully clarified.

Typically, genomic DNA is safeguarded against inappropriate RAG cleavage by the inaccessibility of cryptic RSSs (cRSSs), which are estimated to occur once per 600 base pairs (Lewis, et al., 1997; Teng, et al., 2015). However, recent research has demonstrated that epigenetic reprogramming in cancer can result in heritable alterations in gene expression, including the accessibility of cRSSs (Khoshchehreh, et al., 2019; Becker, et al., 2020; Fatma, et al., 2022; Goel, et al., 2022). We selected the BCRABL1+ B-ALL model, which is characterized by ongoing V(D)J recombinase activity and BCR-ABL1 gene rearrangement in pre-B leukemic cells (Schjerven, et al., 2017; Wong and Witte, 2004). The genome structural variations (SVs) analysis was conducted on leukemic cells from fRAG, cRAG1, and cRAG2, BCR-ABL1+ B-ALL mice to examine the involvement of non-core RAG regions in offtarget V(D)J recombination events. The non-core domain deletion in both RAG1 and RAG2 led to accelerated leukemia onset and progression, as well as an increased off-target V(D)J recombination. Our analysis showed a reduction in RAG binding accuracy in cRAG cells and a decrease in recombinant size in cRAG1 cells, which may be responsible for the increased off-target V(D)J recombination in cRAG leukemia cells. In conclusion, our results highlight the potential importance of the non-core RAG region, particularly RAG1’s non-core region, in maintaining accuracy of V(D)J recombination and genomic stability in BCR-ABL1+ B-ALL.

Method

Mice

The C57BL/6 mice were procured from the Experimental Animal Center of Xi’an Jiaotong University, while cRAG1 (amino acids 384-1040) and cRAG2 (amino acids 1-383) were obtained from David G. Schatz (Yale University, New Haven, Connecticut, USA). The mice were bred and maintained in a specific pathogen-free (SPF) environment at the Experimental Animal Center of Xi’an Jiaotong University. All animal-related procedures were in accordance with the guidelines approved by the Xi’an Jiaotong University Ethics Committee for Animal Experiments.

Generation of Retrovirus Stocks

The pMSCV-BCR-BAL1-IRES-GFP vector is capable of co-expressing the human BCR-ABL1 fusion protein and green fluorescence protein (GFP), while the pMSCV-GFP vector serves as a negative control by solely expressing GFP. To produce viral particles, 293T cells were transfected with either the MSCV-BCRBAL1-IRES-GFP or MSCV-GFP vector, along with the packaging vector PKAT2, utilizing the X-tremeGENE HP DNA Transfection Reagent from Roche (Basel, Switzerland). After 48 hours, the viral supernatants were collected, filtered, and stored at -80°C.

Bone Marrow Transduction and Transplantation

Experiments were conducted using mice aged between 6 to 10 weeks. BCR-ABL1+ B-ALL was induced by utilizing marrow from donors who had not undergone 5-FU treatment. The donor mice were euthanized through CO2 asphyxiation, and the bone marrow was harvested by flushing the femur and tibia with a syringe and 26-gauge needle. Erythrocytes were not removed, and 1 ×106 cells per well were plated in six-well plates. A single round of cosedimentation with retroviral stock was performed in medium containing 5% WEHI-3B-conditioned medium and 10 ng/mL IL-7 (Peprotech, USA). After transduction, cells were either transplanted into syngeneic female recipient mice (1 × 106 cells each) that had been lethally irradiated (2 × 450 cGy), or cultured in RPMI-1640 (Hyclone, Logan, UT) medium supplemented with 10% fetal calf serum (Hyclone), 200 mmol/L L-glutamine, 50 mmol/L 2mercaptoethanol (Sigma, St Louis, MO), and 1.0 mg/ml penicillin/streptomycin (Hyclone). Subsequently, recipient mice were monitored daily for indications of morbidity, weight loss, failure to thrive, and splenomegaly. Weekly assessment of peripheral blood GFP percentage was done using FACS analysis of tail vein blood. Hematopoietic tissues and cells were utilized for histopathology, in vitro culture, FACS analysis, secondary transplantation, genomic DNA preparation, protein lysate preparation, or lineage analysis, contingent upon the unique characteristics of mice under study.

Secondary Transplants

Thawed BM cells were sorted using a BD FACS Aria II (Becton Dickinson, San Jose California, USA). GFP positive leukemic cells (1× 106, 1×105, 1×104, and 1×103) were then resuspended in 0.4 mL Hank’s Balanced Salt Solution (HBSS) and intravenously administered to unirradiated syngeneic mice.

Flow cytometry analysis and sorting

Bone marrow, spleen cells, and peripheral blood were harvested from leukemic mice. Red blood cells were eliminated using NH4Cl RBC lysis buffer, and the remaining nucleated cells were washed with cold PBS. In order to conduct in vitro cell surface receptor staining, 1× 106 cells were subjected to antibody staining for 20 minutes at 4°C in 1×phosphate buffer saline (1×PBS) containing 3% BSA. Cells were then washed with 1×PBS and analyzed using a CytoFLEX Flow Cytometer (Beckman Coulter, Miami, FL) or sorted on a BD FACS Aria II. Apoptosis was analysed by resuspending the cells in Binding Buffer (BD Biosciences, Baltimore, MD, USA), and subsequent labeling with anti-annexin V-AF647 antibody (BD Biosciences) and propidium iodide (BD Biosciences) for 15 minutes at room temperature. The lineage analysis was performed using the following antibodies, which were procured from BD Biosciences: anti-BP-1-PerCP-Cy7, anti-CD19-PerCP-CyTM5.5, antiCD43-PE, anti-B220-APC, and anti-μHC-APC.

BrdU incorporation and analysis

Cells obtained from primary leukemic mice were cultured in sixwell plates containing RPMI-1640 medium supplemented with 10% FBS and 50 mg/ml BrdU. After a 30-minute incubation at 37°C, cells were harvested and intranuclearly stained using anti-BrdU and 7-AAD antibodies, as per the manufacturer’s instructions.

Western blotting analysis

Over 1 × 106 leukemic cells were centrifuged and washed with icecold PBS. The cells were then treated with ice-cold RIPA buffer, consisting of 50 mM Tris-HCl (pH 7.4), 0.15 M NaCl, 1% Triton X-100, 0.5% NaDoc, 0.1% sodium dodecyl sulphide (SDS), 1 mM ethylene diamine tetraacetic acid (EDTA), 1 mM phenylmethane sulphony fluoride (PMSF) (Amresco), and fresh protease inhibitor cocktail Pepstain A (Sigma). After sonication using a Bioruptor TMUCD-200 (Diagenode, Seraing, Belgium), the suspension was spined at 14,000 g for 3 minutes at 4°C. The total cell lysate was either utilized immediately or stored at -80°C. Protein concentrations were determined using DC Protein Assay (Bio–Rad Laboratories, Hercules, California, USA). Subsequently, the protein samples (20 μg) were incubated with α-RAG1 (mAb 23) and α-RAG2 (mAb 39) antibodies, with GAPDH serving as the loading control. The signal was further detected using secondary antibody of goat anti-rabbit IgG conjugated with horseradish peroxidase (Thermo Scientific, Waltham, MA). The band signal was developed with Immobilon™ Western Chemiluminescent HRP substrate (Millipore, Billerica, MA). The band development was analyzed using GELPRO ANALYZER software (Media Cybernetics, Bethesda, MD).

Genomic PCR

Genomic PCR was performed using the following primers (Schlissel, et al., 1991):

DhL-5′-GGAATTCGMTTTTTGTSAAGGGATCTACTACTGTG-3′; J3-5′-GTCTAGATTCTCACAAGAGTCCGATAGACCCTGG-3′; VQ52-5′-CGGTACCAGACTGARCATCASCAAGGACAAYTCC-3′; Vh558-5′-CGAGCTCTCCARCACAGCCTWCATGCARCTCARC-3′; Vh7183-5′-CGGTACCAAGAASAMCCTGTWCCTGCAAATGASC-3′.

RNA-seq library preparation and sequencing

GFP+CD19+ cells were sorted from the spleen of cRAG1 (n=3, 1×106 cells /sample), cRAG2 (n=3, 1×106 cells /sample), and fRAG (n=3, 1×106 cells /sample) B-ALL mice. Total RNA was extracted using Trizol reagent (Invitrogen, CA, USA) following the manufacturer’s guidelines. RNA quantity and purity analysis was done using Bioanalyzer 2100 and RNA 6000 Nano LabChip Kit (Agilent, CA, USA) with RIN number >7.0. RNA-seq libraries were prepared by using 200 ng total RNA with TruSeq RNA sample prep kit (Illumina). Oligo(dT)-enriched mRNAs were fragmented randomly with fragmentation buffer, followed by firstand second-strand cDNA synthesis. After a series of terminal repair, the double-stranded cDNA library was obtained through PCR enrichment and size selection. cDNA libraries were sequenced with the Illumina Hiseq 2000 sequencer (Illumina HiSeq 2000 v4 Single-Read 50 bp) after pooling according to its expected data volume and effective concentration. Two biological replicates were performed in the RNA-seq analysis. Raw reads were then aligned to the mouse genome (GRCm38) using Tophat2 RNA-seq alignment software, and unique reads were retained to quantify gene expression counts from Tophat2 alignment fles. The differentially expressed mRNAs and genes were selected with log2 (fold change) >1 or log2 (fold change) <-1 and with statistical significance (p value < 0.05) by R package. Bioinformatic analysis was performed using the OmicStudio tools athttps://www.omicstudio.cn/tool.

Preparation of tumor DNA samples

GFP+CD19+ splenic cells, tail and kidney tissue were obtained from cRAG1, cRAG2 and fRAG BCR-ABL1+ B-ALL mice, and genomic DNA was extracted using a TIANamp Genomic DNA Kit (TIANGEN-DP304). Subsequently, paired-end libraries were constructed from 1 µg of the initial genomic material using the TruSeq DNA v2 Sample Prep Kit (Illumina, #FC-121-2001) as per the manufacturer’s instructions. The size distribution of the libraries was assessed using an Agilent 2100 Bioanalyzer (Agilent Technologies, #5067-4626), and the DNA concentration was quantified using a Qubit dsDNA HS Assay Kit (Life Technologies, #Q32851). The Illumina HiSeq 4000 was utilized to sequence the samples, with two to four lanes allocated for sequencing the tumor and one lane for the control DNA library of the kidney or liver, each with 150 bp paired end reads.

Read alignment and structural variant calling

Fastq files were generated using Casava 1.8 (Illumina), and BWA 37 was employed to align the reads to mm9. PCR duplicates were eliminated using Picard’s Mark Duplicates tool (sourceforge.net/apps/mediawiki/picard). Our custom scripts (http://sourceforge.net/projects/svdetection) were utilized to eliminate BWA-designated concordant and read pairs with low BWA mapping quality scores. Intrachromosomal and inter-chromosomal rearrangements were identified using SV Detect from discordant, quality prefiltered read pairs. The mean insertion size and standard deviation for this analysis were obtained through Picard’s InsertSizeMetrics tool (sourceforge.net/apps/mediawiki/picard). Tumor-specific structural variants (SVs) were identified using the manta software (https://github.com/Illumina/manta/blob/mater/docs/userGuide/REA DME.md#introduction).

Validation of high confidence off-target candidates

The elimination of non-specific structural mutations from the kidney or tail was necessary for tumor-specific structural variants identification. Subsequently, the method involving 21-bp CAC-tobreakpoint was employed to filter RAG-mediated off-target gene. The validation of high confidence off-target candidates was carried out through PCR. Oligonucleotide primers were designed to hybridize within the “linking” regions of SV Detect, in the appropriate orientation. The PCR product was subjected to Sanger sequencing and aligned to the mouse mm9 reference genome using BLAST (https://blast.ncbi.nlm.nih.gov/Blast.cgi).

Statistics

Statistical analysis was conducted using SPSS 20.0 (IBM Corp.) and GraphPad Prism 6.0 (GraphPad Software). Descriptive statistics were reported as means ± standard deviation for continuous variables. The equality of variances was assessed using Levene’s test. Two-group comparisons, multiple group comparisons, and survival comparisons were performed using independent-samples t-test, one-way analyses of variance (ANOVA) with post hoc Fisher’s LSD test, and log-rank Mantel-Cox analysis, respectively. Kaplan-Meier survival curves were utilized to depict the changes in survival rate over time. Statistical significance was set at P<0.05.

Results

cRAG give more aggressive leukemia in a mouse model of BCR-ABL1+ B-ALL

In order to assess the impact of RAG activity on the clonal evolution of BCR-ABL1+ B-ALL through a genetic experiment, we utilized bone marrow transplantation (BMT) to compare disease progression in fRAG, cRAG1, and cRAG2 BCR-ABL1+ B-ALL (Yu, et al., 2019). Bone marrow cells transduced with a BCR-ABL1/GFP retrovirus were administered into syngeneic lethally irradiated mice, and CD19+ B cell leukemia developed within 30-80 days (Fig. 1A, Fig. S1). Western blot results confirmed equivalent transduction efficiencies of the retroviral BCR-ABL1 in all three cohorts (Fig. S2A). In order to investigate potential variances in leukemia outcome across distinct genomic backgrounds, we used Mantel-Cox estimation to assess survival rates in fRAG, cRAG1, or cRAG2 mice that were transplanted with BCR-ABL1-transformed bone marrow cells. Our findings indicate that, compared to fRAG BCRABL1+ B-ALL mice, cRAG1 or cRAG2 BCR-ABL1+ B-ALL mice exhibited reduced survival rates during the primary transplant phase (median 74.5 days versus 39 or 57 days, P < 0.0425, Fig. 1A). This survival rates discrepancy was also observed during the secondary transplant phase, wherein leukemic cells were extracted from the spleens of primary recipients and subsequently purified via GFP+ cell sorting. A total of 105,104 and 103 GFP+ leukemic cells that originated from fRAG, cRAG1, or cRAG2 leukemic mice were transplanted into corresponding non-irradiated immunocompetent syngenetic recipient mice (median survival days 11-26, 10-16, 11-21 days, P < 0.0023-0.0299, Fig. S2B). Additionally, the cRAG mice exhibited a significantly higher leukemia burden in the bone marrow, spleen, and peripheral blood compared to the fRAG mice (Fig. 1B-D). In order to investigate the cellular process underlying the increased growth rate in cRAG BCRABL1+ B-ALL, a flow cytometry analysis was conducted to examine the cell cycle and cell apoptosis status. The results demonstrated that cRAG BCR-ABL1+ B-ALL had elevated proportion of cells in S/G2-M phase compared to fRAG (Fig. 1E). Furthermore, we observed increased growth due to decreased apoptosis in cRAG leukemic cells (Fig. S2C). RNA-seq analysis revealed the changes of cell differentiation and proliferation/apoptotic pathways (Fig. S3) These results suggest that the absence non-core RAG regions of expedites the progression of malignant transformation and leukemic growth, resulting in aggressive disease phenotype in the cRAG BCR-ABL1+ B-ALL model.

cRAGs give more aggressive leukemia in mice model of BCR-ABL1+ B-ALL (A) Kaplan-Meier survival curve for fRAG (n=8), cRAG1 (n=6), and cRAG2 (n=10) recipient mice. The survival was calculated by Mantel–Cox test (P<0.0425). (B) The spleen weights of fRAG, cRAG1 and cRAG2 leukemic mice (fRAG, n=8, cRAG1, n=7, cRAG2, n=9; fRAG vs cRAG1, P<0.0001, fRAG vs cRAG2, P=0.1352). (C) The spleen cell numbers of fRAG, cRAG1 and cRAG2 leukemic mice (fRAG, n=7, cRAG1, n=8, cRAG2, n=13; fRAG vs cRAG1, P=0.0047, fRAG vs cRAG2, P=0.0180). (D) The percentage of GFP+ cells in peripheral blood (PB, fRAG, n=6, cRAG1, n=6, cRAG2, n =6; fRAG vs cRAG1, P=0.0003, fRAG vs cRAG2, P=0.0035), bone marrow (BM, fRAG, n=5, cRAG1, n=5, cRAG2, n=6; fRAG vs cRAG1, P=0.0341, fRAG vs cRAG2, P=0.0008), and spleen (SP, fRAG, n=9, cRAG1, n=4, cRAG2, n=9; fRAG vs cRAG1, P=0.0016, fRAG vs cRAG2, P<0.0001) of fRAG, cRAG1 and cRAG2 leukemic mice. (E) Representative flow cytometry plots of cell cycle arrest of leukemic cells in fRAG, cRAG1 and cRAG2 mice. In the graph, the percentages of each phase of the cell cycle are summarized below (fRAG, n=3, cRAG1, n=5, cRAG2, n=5; G0/G1, fRAG vs cRAG1, P=0.0082, fRAG vs cRAG2, P=0.0279; S, fRAG vs cRAG1, P=0.0146, fRAG vs cRAG2, P=0.0370; G2/M, fRAG vs cRAG1, P=0.0134, fRAG vs cRAG2, P=0.1507). In figures B, C, D and J, error bars represent the mean ± s.d., P values were calculated by Student’s t test and *P < 0.05, **P < 0.01, ***P < 0.001.

The loss of non-core RAG regions corresponds to a less mature cell surface phenotype but does not impede IgH VDJ recombination

To identify the B cells developmental stages from which the accumulated B leukemic cells originated. We stained single cells with B cell-specific surface markers and analyzed the samples on flow cytometry. We observed that 91%-98% of GFP+ cells from cRAG mice were CD19+BP-1+B220+CD43+, suggesting that the majority of leukemic cells belonged to the large pre-B cell stage. However, in fRAG leukemic mice, there were 65% large pre-B (GFP+CD19+BP-1+B220+CD43+) and 35% small pre-B cells (GFP+CD19+BP-1+B220+CD43-) (Fig. 2A). The lineage results in accordance with the immunoglobulin μ heavy chain (μHC) expression. Specifically, 5% of fRAG leukemic cells exhibited μHC expression, while cRAG leukemic cells lacked μHC expression, indicating a deficiency in the pre-BCR checkpoint (Fig. 2B). These results suggest that leukemic cells derived from cRAG mice arrest at early B cell developmental stage. Typically, IgH rearrangement initiates with D-J joining in pro-B cells, followed by V-DJ joining in large pre-B cells, and ultimately, V-J rearrangements occur at the IgL loci in small pre-B cells. Genomic PCR of GFP+CD19+ cell’s DNA was used to investigate VDJ rearrangement. Notably, cRAG leukemic cells exhibited a high degree of oligoclonality, as all tumors analyzed consistently displayed the rearrangement of a limited number of VH family members. In contrast, the fRAG leukemias exhibited a significant degree of polyclonality, as confirmed by the recurrent rearrangement of multiple VH family members investigated, each rearranged to all possible JH1-3 segments (Fig. S4AB). This finding is consistent with the more aggressive disease phenotype observed in cRAG BCR-ABL1+ B-ALL. The progression of BCR-ABL-induced leukemia in cRAG mice necessitates the selection of secondary oncogenic events, ultimately leading to the emergence of one or a few dominant leukemic clones. Loss of the non-core RAG region results in the emergence of fewer leukemic clones to generate oligoclonal tumors.

The non-core RAG region loss corresponds to a less mature cell surface phenotype (A) Flow cytometry analysis of the B cell markers CD19, BP-1, B220, and CD43 on BCR-ABL1-transformed fRAG, cRAG1 and cRAG2 leukemic bone marrow cells. The percentages of each phase of the B cell stage are summarized in the bottom graph (fRAG, n=9, cRAG1, n=4, cRAG2, n=9; Large-preB, fRAG vs cRAG1, P=0.0349, fRAG vs cRAG2, P=0.0017; Small-pre-B, fRAG vs cRAG1, P=0.0141, fRAG vs cRAG2, P=0.0005). The expression of the cytoplasmic μ chain was analyzed by flow cytometry. Representative samples are shown in (B), and the results from multiple samples analyzed in independent experiments are summarized in the bottom graph as the fraction of cells expressing cytoplasmic factors (fRAG, n=11, cRAG1, n=8, cRAG2, n=8; fRAG vs cRAG1, P=0.3020, fRAG vs cRAG2, P=0.2267). Error bars represent the mean ± s.d., P values were calculated by Student’s t test and *P < 0.05, **P < 0.01, ***P < 0.001.

The loss of non-core RAG regions highlights genomic DNA damage

The aforementioned findings indicate that leukemic cells derived from three kind of mice were hindered in the large pre-B phase to varying extents, instead of consistent development. During this phase, typical B cells show reduced RAG1 expression to accommodate DNA replication and rapid cell proliferation. The non-core RAG regions contain the RING finger domain and T490 residue, which are responsible for regulating RAG degradation in B cell development. Therefore, it is imperative to investigate the potential consequences of non-core region deletion on RAG expression and function in these leukemic cells. Western blotting was done to address this question. The results revealed that RAG1 (cRAG1) and RAG2 (cRAG2) were expressed in GFP+CD19+ splenic leukemic cells derived from BCR-ABL1+ B-ALL mice with varying genetic backgrounds (Fig. 3A). Notably, upregulation of the RAG1 (cRAG1) protein was observed in cRAG1 leukemic cells compared to fRAG (Fig. 3A, Fig. S5A). The in vitro V(D)J recombination assay confirmed that rearrangements mediated by RSS occurred in leukemic cells. This finding suggests that different forms of RAG exhibited cleavage activity (Fig. 3B and Fig. S5B).

The non-core RAG region loss highlights genomic DNA damage (A) Western blotting analysis showed RAG1 and RAG2 expression in GFP+CD19+ leukemic cells originating from BCR-ABL1+ B-ALL in different genetic backgrounds. (B) Rearrangement substrate retrovirus was transduced into leukemic cells. Flow cytometry was used to analyze the percentage of CD90.1 and hCD4 positive cells, and the percentage populations are shown in the bottom graph (fRAG, n=3, cRAG1, n=3, cRAG2, n=3; fRAG vs cRAG1, P=0.0002, fRAG vs cRAG2, P=0.5865). (C) Flow cytometry analysis of LH2AX levels in fRAG, cRAG1 and cRAG2 leukemic cells and the percentage of L-H2AX-positive cell populations shown in the bottom graph (fRAG, n=11, cRAG1, n=8, cRAG2, n=8; fRAG vs cRAG1, P=0.0505, fRAG vs cRAG2, P=0.0094). Error bars represent the mean ± s.d., P values were calculated by Student’s t test and *P < 0.05, **P < 0.01, ***P < 0.001.

To investigate the potential correlation between aberrant RAG activities and increased DNA double-strand breaks (DSBs), we evaluated the levels of phosphorylated H2AX (L-H2AX), a DSB response factor, in fRAG, cRAG1, and cRAG2 leukemic cells (gated on GFP+). This served as a measure of both DNA DSBs and global genomic instability. Our flow cytometry findings revealed that cRAG leukemic cells had increased L-H2AX compared to the fRAG compartment (Fig. 3C). This suggests that cRAG play a more significant role in mediating somatic structural variants in BCR-ABL1+ B lymphocytes. These results indicate that the stalled B-precursors exhibit high expression of RAG endonucleases and increased DNA damage.

Off-target recombination mediated by RAG in BCR-ABL1+ B lymphocytes

Genome-wide sequencing and analysis were performed to compare somatic structural variants (SVs) in BCR-ABL1+ B lymphocytes derived from fRAG, cRAG1, and cRAG2 mice. The leukemic cells were sequenced with an average coverage of 25× (Table. S1). The SVs generated by RAG were screened based on two criteria: the presence of a CAC to the right (or GTG to the left) of both breakpoints, and its occurrence within 21 bp from the breakpoint (Mijušković, et al., 2015). Further elaboration on these criteria can be found in Supplementary Figure 6. Consequently, aberrant V-toV junctions and V to intergenic regions were encompassed in five validated abnormal rearrangements at Ig loci in cRAG leukemic mice (Table. S2). Additionally, seven samples had 24 somatic structural variations, with an average of 3.4 coding region mutations per sample (range of 0-9), which is consistent with the limited number of acquired somatic mutations observed in hematological cancers and childhood malignancies (Fig. 4 and Table S3). The results of the study demonstrate that fRAG cells had low SVs (0-1 per sample), cRAG1 cells exhibited higher SVs (6-9 per sample) while cRAG2 cells had moderate SVs incidence (1-4 per sample) (Fig. 4, Table S3). These findings suggest that cRAG may lead to an elevated off-target recombination, eventually posing a threat to the BCR-ABL1+ B lymphocytes genome.

Structural alterations in BCR-ABL1+ B lymphocytes

(A-C) Circos plot representation of all off-target recombination detected in the genome-wide analyses of fRAG, cRAG1 and cRAG2 leukemic cells. See also Table S3.

Off-target V(D)J recombination characteristics in BCR-ABL1+ B lymphocytes

We further analyzed the characteristics of the identified SVs. Specifically, an assessment of the exon-intron distribution profiles of 42 breakpoints engendered by 24 SVs was executed via genome analysis. The results indicated that while 57% of the breakpoints were situated on the gene body, 43% were enriched within the flanking sequence, majority of which were identified as transcriptional regulatory sequence (Figure 5A). P and N nucleotides are recognized as distinctive characteristics of V(D)J recombination. Consequently, the length of P and N nucleotides remains consistent during RSS-to-RSS and cRSS-to-cRSS recombination.

Overview and characteristics of off-target recombination in BCR-ABL1+ B-ALL leukemic cells from fRAG and cRAG mice (A)Exon-intron distribution profiles of 42 breakpoints generated by 24 SVs. Gene body includes exon (n = 9; 17.3%) and intron (n = 20; 38.5%). Flanking sequence includes 3’UTR (n = 6; 11.5%), 5’UTR (n = 2; 3.8%), promoter (n = 6; 11.5%), and downstream (n = 9; 17.3%). (B) the off-target recombination was filtered and verified by whole genomic sequence and PCR respectively. P nucleotides and N nucleotides of RSS to RSS and cRSS to cRSS were calculated in BCR-ABL1+ B-ALL. (C) Hybrid joint percentage generated by either fRAG, cRAG1 or cRAG2 in BCR-ABL1+ B-ALL. It was 0, 100%, and 93% in fRAG, cRAG1 or cRAG2 leukemic cells respectively. (D)The 24 off-target recombination genes were retrieved by COSMIC Cancer Gene Census (http://cancer.sanger.ac.uk/census/). 0.5 genes and 0.5 cancers gene average sample in fRAG leukemic cells; 8 genes and 4.5 cancer genes average sample in cRAG1 leukemic cells; 3.3 genes and 0.3 cancer genes average sample in cRAG2 leukemic cells.

However, the frequency of P and N sequences was 50%/50% (P/N) in RSS-to-RSS recombination, while it was 4%/8% (P/N) in cRSS-to-cRSS recombination (Fig. 5B). The notably reduced frequency of P and N sequences indicating that off-target sites DNA repair in BCR-ABL1+ B lymphocytes differs from classical V(D)J recombination repair.

The hybrid joints were specifically pronounced in cRAG1 and cRAG2 leukemic cells (100% and 93% respectively), indicating that the non-core regions might be involved in suppressing potentially harmful transposition events (Figure 5C). In order to ascertain the potential impact of non-core RAG region deletion on the occurrence of oncogenic mutations, a comparative analysis of cancer genes was performed across three distinct leukemic cell backgrounds. The results revealed that the number of cancer genes produced in the cRAG1 leukemic cell was significantly higher than the other two. This observation aligns with the manifestation of a most aggressive leukemic phenotype and concurrent alterations in mRNA transcription in cRAG1 BCR-ABL1+ B-ALL mice.

The non-core regions have effects on RAG binding accuracy and off-target recombination size in BCR-ABL1+ B lymphocytes

Sequence logos were used to visually compare the RSS and cRS Ss in Ig loci and non-Ig loci respectively. The RSS elements in Ig lo ci exhibited the closest match to the canonical RSS (CACAGTG [1 2/23 spacer] ACAAAAACC), particularly at functionally important p ositions. It is noteworthy that the first 5 bases (underlined) of the p erfect heptamer sequence CACAGTG served as the binding motif of fRAG, while the first 4 bases of the heptamer sequence, the CA CA tetranucleotide, were identified as the cRAG binding motif in le ukemic cells (Fig. 6A). Although both motifs (CACAG and CACA) c orresponding to the highly conserved portion of the RSS heptamer sequence, variations in the cRSSs sequence among off-target gen es in fRAG and cRAG mice indicate that the removal of RAG’s non -core region reduces binding precision and increases the off-target recombination in BCR-ABL1+ B lymphocytes.

The non-core regions have effects on RAG binding accuracy and recombinat size in BCR-ABL1+B lymphocytes (A) Sequence logos were used to compare the RSS and cRSS in Ig loci and non-Ig loci. Top panel: V(D)J recombination at Ig locus; the next three panels: RAG-mediated off-target recombination at non-Ig locus from fRAG, cRAG1 and cRAG2 leukemic cells respectively. The scale of recombinant size was categorized into three ranges: <1000bp, 1000-10000bp, and >10000bp. The distribution of different recombinant sizes in fRAG, cRAG1, and cRAG2 leukemic cells was presented in (B), while the number of different recombinant sizes in fRAG, cRAG1, and cRAG2 leukemic cells was displayed in (C). (D) A schematic depiction of the mechanism of cRAG-accelerated off-target V(D)J recombination was provided. Both RAG1 and RAG2’s non-core region deletion decreases RAG binding accuracy in cRAG1 and cRAG2, BCR-ABL1+ B ALL. Additionally, RAG1’s non-core region deletion significantly reduces the size and scale of off-target V(D)J recombination in cRAG1, BCR-ABL1+ B ALL.

Antigen receptor genes are assembled by large-scale deletions and inversions. Our investigation revealed that both fRAG and cR AG2 leukemic cells generated 100% and 92% off-target recombina tion, respectively, exceeding 10000 bp. However, cRAG1 leukemic cells exhibited only 6% off-target recombination exceeding 10000bp, with 48% being <1000bp, 46% being 1000-10000bp (Fig. 6BC). The results indicate that cRAG1 generate minor size of off-target r ecombination in BCR-ABL1+ B lymphocytes, and non-core RAG1 r egion influences the off-target recombination magnitude. Additional ly, non-core RAG1 region deletion results in reduced off-target reco mbination size, which may account for the higher incidence of off-t arget V(D)J recombination in cRAG1 leukemic cells (Fig. 6D).

Discussion

In this study, we have demonstrated that non-core region deletion of both RAG1 and RAG2 leads to accelerated development of leukemia and increased off-target V(D)J recombination in BCR-ABL1+ B lymphocytes. Furthermore, we report reduced cRAG binding accuracy and off-target recombination size in cRAG1 leukemia cells, which might contribute to exacerbated off-target V(D)J recombination of cRAG BCR-ABL1+ B lymphocytes. These findings suggest that the non-core regions, particularly the non-core region of RAG1, play a crucial role in maintaining accuracy of V(D)J recombination and genomic stability in BCR-ABL1+ B lymphocytes.

Our observations indicate that cRAGs leukemic cells exhibit a heightened production of hybrid joints, and that the non-core RAG regions might suppress hybrid joint generation in vivo. Post-cleavage synaptic complexes (PSCs) consist of the RAGs, coding ends, and RSS ends (Fugmann, et al., 2000; Libri, et al., 2021). It is likely that RAG evolution has resulted in the formation of PSCs with optimal conformation and/or stability for standard coding and RSS end-joining. Conversely, cRAGs PSCs may facilitate RAGmediated hybrid joints by enabling the close proximity of coding and RS ends or increasing the PSC stability. Furthermore, it is possible for fRAGs to enlist disassembly/remodeling factors to PSCs, a process that may facilitate the involvement of NHEJ factors in the completion of the normal reaction (Fugmann, et al., 2000). Within this context, cRAGs might exhibit reduced recruitment capacity due to alterations in overall conformation or the absence of specific motifs, resulting in the formation of more stable PSCs and an increased potential for the accumulation of incomplete hybrid joints. However, our results showed that over 90% junction were hybrid joints in cRAGs leukemic cells, surpassing the frequency previously reported in literature (Raghavan, et al., 2006; Talukder, et al., 2004). Studies have indicated that deficiency in non-homologous end joining (NHEJ) may lead to chromosomal instability and lymphomagenesis (Wiegmans, et al., 2021; Gaymes, et al., 2002; Rassool, 2003; Scully, et al., 2019). Notably, our findings have revealed significant variations in the NHEJ repair path-way among leukemic cells with different genetic backgrounds, suggesting aberrant expression of DNA repair pathways in cRAGs leukemic cells (Fig. S3B). This observation suggests the potential for cRAGs to generate elevated levels of hybrid joints, particularly in the absence of a competing normal pathway for efficient formation of coding and RSS joins in a NHEJ-aberrant background.

The off-target V(D)J recombination process mediated by RAG has the potential to generate oncogenic rearrangements (Mijušković, et al., 2015; Greaves, 2018; Thomson, et al., 2020). Our study reveals that non-core RAG regions deletion increased the off-target oncogenic genes, particularly in cRAG1 leukemic cells. These genes have the capacity to influence cell proliferation, differentiation, or survival. In line with this, the leukemic cells exhibit augmented cell proliferation and reduced cell apoptosis with alterations in the cell cycle pathway, particularly in cRAG1 leukemic cells (Fig. S3DF). CDKN2B is susceptible to recurrent breakage in cRAG1 leukemic cells, and its function in impeding proliferation and enhancing leukemic cells apoptosis is attributed to its capacity to block CDK6 (Lopes-Ventura, et al., 2019; Suzuki, et al., 1995). CDKN2B reduction and CDK6 elevation in cRAG1 leukemic cells have been verified (Fig. S7AB), thereby suggesting that off-target V(D)J recombination generates known or suspected oncogenic mutations. Nevertheless, the degree to which RAG-mediated oncogenic recombination contributes to leukemia necessitates further examination.

In human ETV6-RUNX1 ALL, the ETV6-RUNX1 fusion gene is believed to initiate prenatally, yet the disease remains clinically latent until critical secondary events occur, leading to leukemic transformation-“pre-leukemia to leukemia” (Mori, et al., 2002; Bateman, et al., 2010; Bhojwani, et al., 2012). Genomic rearrangement, mediated by aberrant RAG recombinase activity, is a frequent driver of these secondary events in ETV6-RUNX1 ALL (Papaemmanuil, et al., 2014). In contrast, RAG mediated off-target V(D)J recombination is also observed in BCR-ABL1+ B-ALL. These oncogenic structural variations can also be considered as secondary events that promote the transition -“leukemia to aggressive leukemia”. The enhancement of BCR-ABL1+ B-ALL deterioration and progression by cRAG in mouse model was consistent with our previous study that RAG enhances BCR-ABL1 positive leukemic cell growth through its endonuclease activity. Additionally, we showed that non-core RAG1 region deletion leads to increased cRAG1 expression and high RAG expression related to low survival in pediatric acute lymphoid leukemia (Fig. 3A and Fig. S8). Therefore, more attention should be be paid to the non-core RAG region mutation in BCR-ABL1+ B-ALL for the role of non-core region in leukemia suppression and off-target V(D)J recombination.

Disclosure of Potential Conflicts of Interest

The authors declare no potential conflicts of interest.

Authors’ Contributions

Yanghong Ji: Conceptualization, resources, data curation, funding acquisition, validation, writing-review, and editing. Xiaozhuo Yu and Wen Zhou: Conceptualization, validation, visualization, methodology, writing-original draft, writing-review, and editing. Xiaodong Chen: validation, writing-review, and editing. Shunyu He: methodology, writing-review, and editing. Mengting Qin: writing-review, and editing. Meng Yuan: validation, writing-review, and editing. Yang Wang: validation, writing-review and editing. Woodvine otieno Odhiambo: writing-review and editing. YinSha Miao: funding, validation, writing-review, and editing.

Supporting information

eLife Assessment

The authors have shown a valuable tumor suppressive function of the non-core regions of RAG1/2 recombinases, by using a set of animal models. The work is solid and the conclusions are supported by their data. Some areas of mechanistic work can be improved.

Reviewer #1 (Public Review):

Summary:

In this report, Yu et al ascribe potential tumor suppressive functions to the non-core regions of RAG1/2 recombinases. Using a well-established BCR-ABL oncogene-driven system, the authors model the development of B cell acute lymphoblastic leukemia in mice and found that RAG mutants lacking non-core regions show accelerated leukemogenesis. They further report that the loss of non-core regions of RAG1/2 increases genomic instability, possibly caused by increased off-target recombination of aberrant RAG-induced breaks. The authors conclude that the non-core regions of RAG1 in particular not only increase the fidelity of VDJ recombination, but may also influence the recombination "range" of off-target joints, and that in the absence of the non-core regions, mutant RAG1/2 (termed cRAGs) catalyze high levels of off-target recombination leading to the development of aggressive leukemia.

Strengths:

The authors used a genetically defined oncogene-driven model to study the effect of RAG non-core regions on leukemogenesis. The animal studies were well performed and generally included a good number of mice. Therefore, the finding that cRAG expression led to the development of more aggressive BCR-ABL+ leukemia compared to fRAG is solid.

Weaknesses:

In general, I find the mechanistic explanation offered by the authors to explain how the non-core regions of RAG1/2 suppress leukemogenesis to be less convincing. My main concern is that cRAG1 and cRAG2 are overexpressed relative to fRAG1/2. This raises the possibility that the observed increased aggressiveness of cRAG tumors compared to fRAG tumors could be solely due to cRAG1/2 overexpression, rather than any intrinsic differences in the activity of cRAG1/2 vs fRAG1/2; and indeed, the authors allude to this possibility in Fig S8, where it was shown that elevated expression of RAG (i.e. fRAG) correlated with decreased survival in pediatric ALL. Although it doesn't mean the authors' assertions are incorrect, this potential caveat should nevertheless be discussed.

Some of the conclusions drawn were not supported by the data.

1. I'm not sure that the authors can conclude based on μHC expression that there is a loss of pre-BCR checkpoint in cRAG tumors. In fact, Fig. 2B showed that the differences are not statistically significant overall, and more importantly, μHC expression should be detectable in small pre-B cells (CD43-). This is also corroborated by the authors' analysis of VDJ rearrangements, showing that it has occurred at the H chain locus in cRAG cells.

2. The authors found a high degree of polyclonal VDJ rearrangements in fRAG tumor cells but a much more limited oligoclonal VDJ repertoire in cRAG tumors. They concluded that this explains why cRAG tumors are more aggressive because BCR-ABL induced leukemia requires secondary oncogenic hits, resulting in the outgrowth of a few dominant clones (Page 19, lines 381-398). I'm not sure this is necessarily a causal relationship since we don't know if the oligoclonality of cRAG tumors is due to selection based on oncogenic potential or if it may actually reflect a more restricted usage of different VDJ gene segments during rearrangement.

3. What constitutes a cancer gene can be highly context- and tissue-dependent. Given that there is no additional information on how any putative cancer gene was disrupted (e.g., truncation of regulatory or coding regions), it is not possible to infer whether increased off-target cRAG activity really directly contributed to the increased aggressiveness of leukemia.

4. Fig. 6A, it seems that it is really the first four nucleotide (CACA) that determines fRAG binding and the first three (CAC) that determine cRAG binding, as opposed to five for fRAG and four for cRAG, as the author wrote (page 24, lines 493-497).

5. Fig S3B, I don't really see why "significant variations in NHEJ" would necessarily equate "aberrant expression of DNA repair pathways in cRAG leukemic cells". This is purely speculative. Since it has been reported previously that alt-EJ/MMEJ can join off target RAG breaks, do the authors detect high levels of microhomology usage at break points in cRAG tumors?

6. Fig. S7, CDKN2B inhibits CDK4/6 activation by cyclin D, but I don't think it has been shown to regulate CDK6 mRNA expression. The increase in CDK6 mRNA likely just reflects a more proliferative tumor but may have nothing to do with CDKN2B deletion in cRAG1 tumors.

Insufficient details in some figures. For instance, Fig. 1A, please include statistics in the plot showing a comparison of fRAG vs cRAG1, fRAG vs cRAG2, cRAG1 vs cRAG2. As of now, there's a single p-value (0.0425) stated in the main text and the legend but why is there only one p-value when fRAG is compared to cRAG1 or cRAG2? Similarly, the authors wrote "median survival days 11-26, 10-16, 11-21 days, P < 0.0023-0.0299, Fig. S2B." However, it is difficult for me to figure out what are the numbers referring to. For instance, is 11-26 referring to median survival of fRAG inoculated with three different concentrations of GFP+ leukemic cells or is 11-26 referring to median survival of fRAG, cRAG1, cRAG2 inoculated with 10^5 cells? It would be much clearer if the authors can provide the numbers for each pair-wise comparison, if not in the main text, then at least in the figure legend. In Fig. 5A-B, do the plots depict SVs in cRAG tumors or both cRAG and fRAG cells? Also in Fig. 5, why did 24 SVs give rise to 42 breakpoints, and not 48? Doesn't it take 2 breaks to accomplish rearrangement? In Fig. 6B-C, it is not clear how the recombination sizes were calculated. In the examples shown in Fig. 4, only cRAG1 tumors show intra-chromosomal joins (chr 12), while fRAG and cRAG2 tumors show exclusively inter-chromosomal joins.

Insufficient details on certain reagents/methods. For instance, are the cRAG1/2 mice of the same genetic background as fRAG mice (C57BL/6 WT)? On Page 23, line 481, what is a cancer gene? How are they defined? In Fig. 3C, are the FACS plots gated on intact cells? Since apoptotic cells show high levels of gH2AX, I'm surprised that the fraction of gH2AX+ cells is so much lower in fRAG tumors compared to cRAG tumors. The in vitro VDJ assay shown in Fig 3B is not described in the Method section (although it is described in Fig S5b). Fig. 5A-B, do the plots depict SVs in cRAG tumors or both cRAG and fRAG cells?

Reviewer #2 (Public Review):

Summary: In the manuscript, the authors summarized and introduced the correlation between the non-core regions of RAG1 and RAG2 in BCR-ABL1+acute B lymphoblastic leukemia and off-target recombination which has certain innovative and clinical significance.

Author response:

The following is the authors’ response to the original reviews.

We would like to thank all of the reviewers for their helpful and the effort they made in reading and evaluating our manuscript. In response to them, we have made major changes to the text and figures and performed substantial new experiments. These new data and changes to the text and figures have substantially strengthened the manuscript. We believe that the manuscript is now very strong in both its impact and scope and we hope that reviewers will find it suitable for publication in eLife

A point-by-point response to the reviewers' specific comments is provided below.

Public Reviews:

Reviewer #1 (Public Review):

Summary:

In this report, Yu et al ascribe potential tumor suppressive functions to the non-core regions of RAG1/2 recombinases. Using a well-established BCR-ABL oncogene-driven system, the authors model the development of B cell acute lymphoblastic leukemia in mice and found that RAG mutants lacking non-core regions show accelerated leukemogenesis. They further report that the loss of non-core regions of RAG1/2 increases genomic instability, possibly caused by increased off-target recombination of aberrant RAG-induced breaks. The authors conclude that the non-core regions of RAG1 in particular not only increase the fidelity of VDJ recombination, but may also influence the recombination "range" of off-target joints, and that in the absence of the non-core regions, mutant RAG1/2 (termed cRAGs) catalyze high levels of off-target recombination leading to the development of aggressive leukemia.

Strengths:

The authors used a genetically defined oncogene-driven model to study the effect of RAG non-core regions on leukemogenesis. The animal studies were well performed and generally included a good number of mice. Therefore, the finding that cRAG expression led to the development of more aggressive BCR-ABL+ leukemia compared to fRAG is solid.

Weaknesses:

In general, I find the mechanistic explanation offered by the authors to explain how the non-core regions of RAG1/2 suppress leukemogenesis to be less convincing. My main concern is that cRAG1 and cRAG2 are overexpressed relative to fRAG1/2. This raises the possibility that the observed increased aggressiveness of cRAG tumors compared to fRAG tumors could be solely due to cRAG1/2 overexpression, rather than any intrinsic differences in the activity of cRAG1/2 vs fRAG1/2; and indeed, the authors allude to this possibility in Fig S8, where it was shown that elevated expression of RAG (i.e. fRAG) correlated with decreased survival in pediatric ALL. Although it doesn't mean the authors' assertions are incorrect, this potential caveat should nevertheless be discussed.

We appreciate the valuable suggestions from the reviewer. BCR-ABL1+ B-ALL is characterized by halted early B-lineage differentiation. In BCR-ABL1+ B cells, RAG recombinases are highly expressed, leading to the inactivation of genes that encode essential transcription factors for B-lineage differentiation. This results in cells being trapped within the precursor compartment, thereby elevating RAG gene expression. Our interpretation of the data suggests that, in BCR-ABL1+ B-ALL mouse models, the high expression of both cRAG and fRAG and the deletion of the non-core regions influence the precision of RAG targeting within the genome. This causes more genomic damage in cRAG tumors than in fRAG tumors, consequently leading to the observed increased aggressiveness of cRAG tumors compared to fRAG tumors. We discussed the issues on Page 12, lines 295-307 in the revised manuscript.

Some of the conclusions drawn were not supported by the data.

(1) I'm not sure that the authors can conclude based on μHC expression that there is a loss of pre-BCR checkpoint in cRAG tumors. In fact, Fig. 2B showed that the differences are not statistically significant overall, and more importantly, μHC expression should be detectable in small pre-B cells (CD43-). This is also corroborated by the authors' analysis of VDJ rearrangements, showing that it has occurred at the H chain locus in cRAG cells.

We appreciate the insightful comment from the reviewer. Upon reevaluation of the data presented in Fig. 2B, we identified and rectified certain errors. The revised analysis now shows that the differences in μHC expression are statistically significant. This significant expression of μHC in fRAG leukemic cells implies that these cells may progress further in differentiation, potentially acquiring an immune phenotype. These modifications have been incorporated into the manuscript on page 7, lines 153-156 in the revised manuscript.

(2) The authors found a high degree of polyclonal VDJ rearrangements in fRAG tumor cells but a much more limited oligoclonal VDJ repertoire in cRAG tumors. They concluded that this explains why cRAG tumors are more aggressive because BCR-ABL induced leukemia requires secondary oncogenic hits, resulting in the outgrowth of a few dominant clones (Page 19, lines 381-398). I'm not sure this is necessarily a causal relationship since we don't know if the oligoclonality of cRAG tumors is due to selection based on oncogenic potential or if it may actually reflect a more restricted usage of different VDJ gene segments during rearrangement.

Thank you for your insightful comments and questions regarding the relationship between the oligoclonality of V(D)J rearrangements and the aggressiveness of cRAG tumors. You raise an important point regarding whether the observed oligoclonality is a result of selective pressure favoring clones with specific oncogenic potential, or if it reflects inherent limitations in V(D)J segment usage during rearrangement in cRAG models. In our study, we observed a marked difference in the V(D)J rearrangement patterns between fRAG and cRAG tumor cells, with cRAG tumors exhibiting a more limited, oligoclonal repertoire. This observation led us to speculate that the aggressive nature of cRAG tumors might be linked to a selective advantage conferred by specific V(D)J rearrangements that cooperate with the BCR-ABL1 oncogene to drive leukemogenesis. However, we acknowledge that our current data do not definitively establish a causal relationship between oligoclonality and tumor aggressiveness. The restricted V(D)J repertoire in cRAG tumors could indeed be due to a more constrained rearrangement process, possibly influenced by the altered expression or function of RAG1/2 in the absence of non-core regions. This could limit the diversity of V(D)J rearrangements, leading to the emergence of a few dominant clones not necessarily because they have greater oncogenic potential, but because of a narrowed field of rearrangement possibilities.

To address this question more thoroughly, future studies could examine the functional consequences of specific V(D)J rearrangements found in dominant cRAG tumor clones. This could include assessing the oncogenic potential of these rearrangements in isolation and in cooperation with BCR-ABL1, as well as exploring the mechanistic basis for the restricted V(D)J repertoire. Such studies would provide deeper insight into the interplay between RAG-mediated recombination, clonal selection, and leukemogenesis in BCR-ABL1+ B-ALL.

We appreciate your feedback on this matter and agree that further investigation is required to unravel the precise relationship between V(D)J rearrangement diversity and leukemic progression in cRAG models. We have revised our discussion to reflect these considerations and to clarify the speculative nature of our conclusions regarding the link between oligoclonality and tumor aggressiveness. We added more discussion on this issue on Page 7, lines 166-170 in the revised manuscript.

(3) What constitutes a cancer gene can be highly context- and tissue-dependent. Given that there is no additional information on how any putative cancer gene was disrupted (e.g., truncation of regulatory or coding regions), it is not possible to infer whether increased off-target cRAG activity really directly contributed to the increased aggressiveness of leukemia.

We totally agree you raised the issues. In Supplementary Table 3, we have presented data on off-target gene disruptions, specifically in introns, exons, downstream regions, promoters, 3' UTRs, and 5' UTRs. However, this dataset alone does not suffice to conclusively determine whether the increased off-target activity of cRAG directly influences the heightened aggressiveness of leukemia. To bridge this knowledge gap, our future research will extend to include both knockout and overexpression experiments targeting these off-target genes.

(4) Fig. 6A, it seems that it is really the first four nucleotide (CACA) that determines fRAG binding and the first three (CAC) that determine cRAG binding, as opposed to five for fRAG and four for cRAG, as the author wrote (page 24, lines 493-497).

We thank the reviewer for the insightful comment. In response, we have revised the text to accurately reflect the nucleotide sequences responsible for RAG binding and cleavage. Specifically, we now clarify that the first four nucleotides (CACA) are crucial for fRAG binding and cleavage, while the initial three nucleotides (CAC) are essential for cRAG binding and cleavage. These updates have been made on page 10, lines 242-245 of the revised manuscript.

(5) Fig S3B, I don't really see why "significant variations in NHEJ" would necessarily equate "aberrant expression of DNA repair pathways in cRAG leukemic cells". This is purely speculative. Since it has been reported previously that alt-EJ/MMEJ can join off target RAG breaks, do the authors detect high levels of microhomology usage at break points in cRAG tumors?

We appreciate the reviewer's comment. Currently, we have not observed microhomology usage at breakpoints in cRAG tumors. We plan to address this aspect in a future, more detailed study. Regarding the 'aberrant expression of DNA repair pathways in cRAG leukemic cells, we acknowledge that this is speculative. Therefore, we have carefully rephrased this to 'suggesting a potential aberrant expression of DNA repair pathways in cRAG leukemic cells.' This modification is reflected on page 12, lines 290-291 of the revised manuscript.

(6) Fig. S7, CDKN2B inhibits CDK4/6 activation by cyclin D, but I don't think it has been shown to regulate CDK6 mRNA expression. The increase in CDK6 mRNA likely just reflects a more proliferative tumor but may have nothing to do with CDKN2B deletion in cRAG1 tumors.

We fully concur with the reviewer's comment. We have deleted this inappropriate part from the text.

Insufficient details in some figures. For instance, Fig. 1A, please include statistics in the plot showing a comparison of fRAG vs cRAG1, fRAG vs cRAG2, cRAG1 vs cRAG2. As of now, there's a single p-value (0.0425) stated in the main text and the legend but why is there only one p-value when fRAG is compared to cRAG1 or cRAG2? Similarly, the authors wrote "median survival days 11-26, 10-16, 11-21 days, P < 0.0023-0.0299, Fig. S2B." However, it is difficult for me to figure out what are the numbers referring to. For instance, is 11-26 referring to median survival of fRAG inoculated with three different concentrations of GFP+ leukemic cells or is 11-26 referring to median survival of fRAG, cRAG1, cRAG2 inoculated with 10^5 cells? It would be much clearer if the authors can provide the numbers for each pair-wise comparison, if not in the main text, then at least in the figure legend. In Fig. 5A-B, do the plots depict SVs in cRAG tumors or both cRAG and fRAG cells? Also in Fig. 5, why did 24 SVs give rise to 42 breakpoints, and not 48? Doesn't it take 2 breaks to accomplish rearrangement? In Fig. 6B-C, it is not clear how the recombination sizes were calculated. In the examples shown in Fig. 4, only cRAG1 tumors show intra-chromosomal joins (chr 12), while fRAG and cRAG2 tumors show exclusively inter-chromosomal joins.

We appreciate the reviewer's feedback and have made the following revisions:

(1) The text has been adjusted to rectify the previously mentioned error in the figure legends (page 1, lines 5-6).

(2) We have clarified the intended message in the revised text (page 6, lines 129-130) and the figure legend (page 4-5, lines 107-113) for greater precision.

(3) Figure 5A-B now presents an overview of all structural variants (SVs) identified in both cRAG and fRAG cells, offering a comprehensive comparison.

(4) Among the analyzed SVs, 24 generated a total of 48 breakpoints, with 41 occurring within gene bodies and the remaining 7 in adjacent flanking sequences. This informs our exon-intron distribution profile analysis.

(5) We have defined recombination sizes as ‘the DNA fragment size spanning the two breakpoints’ for clarity (page 10, lines 251-252).

(6) All off-target recombinations identified in the genome-wide analyses of fRAG, cRAG1, and cRAG2 leukemic cells were determined to be intra-chromosomal joins, highlighting their specific nature within the genomic context.

Insufficient details on certain reagents/methods. For instance, are the cRAG1/2 mice of the same genetic background as fRAG mice (C57BL/6 WT)? On Page 23, line 481, what is a cancer gene? How are they defined? In Fig. 3C, are the FACS plots gated on intact cells? Since apoptotic cells show high levels of gH2AX, I'm surprised that the fraction of gH2AX+ cells is so much lower in fRAG tumors compared to cRAG tumors. The in vitro VDJ assay shown in Fig 3B is not described in the Method section (although it is described in Fig S5b). Fig. 5A-B, do the plots depict SVs in cRAG tumors or both cRAG and fRAG cells?

We are grateful for the reviewer's feedback and have incorporated their insights as follows:

(1) We clarify that both cRAG1/2 and fRAG mice share the same genetic background, specifically the C57BL/6 WT strain, ensuring consistency across experimental models.

(2) We define a 'cancer gene' as one harboring somatic mutations implicated in cancer. To support our analysis, we refer to the Catalogue Of Somatic Mutations In Cancer (COSMIC) at http://cancer.sanger.ac.uk/cosmic. COSMIC serves as the most extensive repository for understanding the role of somatic mutations in human cancers.

(3) Upon thorough review of the raw data for γ-H2AX and the fluorescence-activated cell sorting (FACS) plots gated on intact cells, we propose that the observed discrepancies might stem from the limited sensitivity of the γ-H2AX flow cytometry detection method. This insight prompts our commitment to employing more efficient detection methodologies in forthcoming studies.

(4) Detailed procedures for the in vitro V(D)J recombination assay have been included in the Methods section (page 15, lines 384-388) to enhance the manuscript's comprehensiveness and reproducibility.

(5) The presented plots offer a comprehensive overview of structural variants (SVs) identified in both cRAG and fRAG cells, providing a holistic view of the genomic landscape across different models.

Reviewer #3 (Public Review):

Summary:

In the manuscript, the authors summarized and introduced the correlation between the non-core regions of RAG1 and RAG2 in BCR-ABL1+acute B lymphoblastic leukemia and off-target recombination which has certain innovative and clinical significance.

Recommendations For The Authors:

Reviewer #1 (Recommendations For The Authors):

I would suggest that the authors tone down some of their conclusions, which are not necessarily supported by their own data. in addition, there are some minor mistakes in figure assembly/presentation. For instance, I believe that the axes labels in Fig. 1E were flipped. BrdU should be on y-axis and 7-AAD on the x-axis. Fig. 3B, the y-axis contains a typo, it should be "CD90.1..." and not "D90.1...". In Fig. 5C, the numbers seem to be flipped, with 93% corresponding to cRAG1 and 100% to cRAG2 (compare with the description on page 23, lines 474-475). Fig. 5C, y-axis, "hybrid" is a typo. Page 3, line 59: The abbreviation of RSS has already been described earlier (p4, line 53).

We thank the reviewer for these suggestions. We carefully checked the raw data and corrected these mistakes in the revised manuscript.

Page 3, line 63: "signal" segment (commonly referred to as signal ends), not "signaling" segment.

We have changed “signaling segment” to “signal ends in the revised manuscript. (page 3, lines 54-55)

Page 3, lines 64-65: VDJ recombination promotes the development of both B and T cells, and aberrant recombination can cause both B and T cell lymphomas.

The statement about the role of V(D)J recombination in B and T cell development and its link to lymphomagenesis is grounded in a substantial body of research. Theoretical frameworks and empirical studies delineate how aberrations in the recombination process can lead to genomic instability, potentially triggering oncogenic events. This connection is extensively documented in immunology and oncology literature, illustrating the critical balance between necessary genetic rearrangements for immune diversity and the risk of malignancy when these processes are dysregulated (Thomson, et al.,2020; Mendes, et al.,2014; Onozawa and Aplan,2012).

Page 4, line 72: "recombinant dispensability" is not a commonly used phrase. Do the authors mean the say that the non-core regions of RAG1/2 are not strictly required for VDJ recombination?

We thank the reviewers for their insightful suggestion. We have revised the sentence to read, 'Although the non-core regions of RAG1/2 are not essential for V(D)J recombination, the evolutionary conservation of these regions suggests their potential significance in vivo, possibly affecting RAG activity and expression in both quantitative and qualitative manners.' This revision appears on page 3, lines 61-62, in the revised manuscript.

Fig. 4. It would have been nice to show at least one more cRAG1 tumor circus plot.

We appreciate the reviewer's comment and concur with the suggestion. In future sequencing experiments, we will consider including additional replicates. However, due to time and financial constraints, the current sequencing effort was limited to a maximum of three replicates.

Reviewer #3 (Recommendations For The Authors):

In the manuscript, the authors summarized and introduced the correlation between the non-core regions of RAG1 and RAG2 in BCR-ABL1+acute B lymphoblastic leukemia and off-target recombination which has certain innovative and clinical significance. The following issues need to be addressed by the authors.

(1) Authors should check and review extensively for improvements to the use of English.

We thank the reviewer for their comment. With assistance from a native English speaker, we have carefully revised the manuscript to enhance its readability.

(2) Authors should revise the conclusion so that the above can be clearly reviewed and summarized.

The conclusion has been partially revised in the revised manuscript.

(3) The article should state that the experiment was independently repeated three times.

The experiment was repeated under the same conditions three times and the information has been descripted in Statistics section on page 19, lines 473-475 in the revised manuscript.

(4) The article will be more convincing if it uses references in the last 5 years.

We are grateful to the reviewer for their guidance in enhancing our manuscript. We have incorporated additional references from the past five years in the revised version.

(5) Additional experiments are suggested to elucidate the molecular mechanisms related to off-target recombination.

We thank the reviewer for this suggestion. In future experiments, we plan to perform ChIP-seq analysis to investigate the relationship between chromatin accessibility and off-target effects, as well as to examine the impact of knocking out and overexpressing off-target genes on cancer development and progression.

(6) It is suggested to further analyze the effect of the absence of non-core RAG region on the differentiation and development of peripheral B cells in mice by flow analysis and expression of B1 and B2.

Thank you very much for highlighting this crucial issue. FACS analysis was performed, revealing that leukemia cells in peripheral B cells in mice did not express CD5. The data are presented as follows:

(7) Fig3A should have three biological replicates and the molecular weight should be labeled on the right side of the strip.

Thank you for this suggestion. The experiment was independently repeated three times, and the molecular weights have been labeled on the right side of the bands in the revised version

References:

Mendes RD, Sarmento LM, Canté-Barrett K, Zuurbier L, Buijs-Gladdines JG, Póvoa V, Smits WK, Abecasis M, Yunes JA, Sonneveld E, Horstmann MA, Pieters R, Barata JT, Meijerink JP. 2014. PTEN microdeletions in T-cell acute lymphoblastic leukemia are caused by illegitimate RAG-mediated recombination events. BLOOD 124:567-578. doi:10.1182/blood-2014-03-562751

Onozawa M, Aplan PD. 2012. Illegitimate V(D)J recombination involving nonantigen receptor loci in lymphoid malignancy. Genes Chromosomes Cancer 51:525-535. doi:10.1002/gcc.21942

Thomson DW, Shahrin NH, Wang P, Wadham C, Shanmuganathan N, Scott HS, Dinger ME, Hughes TP, Schreiber AW, Branford S. 2020. Aberrant RAG-mediated recombination contributes to multiple structural rearrangements in lymphoid blast crisis of chronic myeloid leukemia. LEUKEMIA 34:2051-2063. doi:10.1038/s41375-020-0751-y

Version 2

Abstract

Abstract

The evolutionary conservation of non-core RAG regions suggests significant roles that might involve quantitative or qualitative alterations in RAG activity. Off-target V(D)J recombination contributes to lymphomagenesis and is exacerbated by RAG2’ C-terminus absence in Tp53-/- mice thymic lymphomas. However, the genomic stability effects of non-core regions from both cRAG1 and cRAG2 in BCR-ABL1+ Blymphoblastic leukemia (BCR-ABL1+ B-ALL), the characteristics, and mechanisms of non-core regions in suppressing off-target V(D)J recombination remains unclear. Here, we established three mouse models of BCR-ABL1+ B-ALL in mice expressing full-length RAG (fRAG), core RAG1 (cRAG1), and core RAG2 (cRAG2). The cRAG (cRAG1 and cRAG2) leukemia cells exhibited greater malignant tumor characteristics compared to fRAG cells. Additionally, cRAG cells showed higher frequency of off-target V(D)J recombination and oncogenic mutations than fRAG. We also revealed decreased RAG cleavage accuracy in cRAG cells and a smaller recombinant size in cRAG1 cells, which could potentially exacerbate off-target V(D)J recombination in cRAG cells. In conclusion, these findings indicate that the non-core RAG regions, particularly the non-core region of RAG1, play a significant role in preserving V(D)J recombination precision and genomic stability in BCR-ABL1+ B-ALL.

Introduction

V(D)J recombination serves as the central process for early lymphocyte development and generates diversity in antigen receptors. This process involves the double-strand DNA breaks of gene segments by the V(D)J recombinase, including RAG1 and RAG2. RAG recognizes conserved recombination signal sequences (RSSs) positioned adjacent to V, D, and J gene segments. A bona fide RSS contains a conserved palindromic heptamer (consensus 5′-CACAGTG) and A-rich nonamer (consensus 5′-ACAAAAACC) separated by a degenerate spacer of either 12 or 23 base pairs (Hirokawa, et al., 2020; Schatz and Ji, 2011). The process of efficient recombination is contingent upon the presence of recombination signal sequences (RSSs) with differing spacer lengths, as dictated by the “12/23 rule” (Banerjee and Schatz, 2014;Eastman, et al., 1996;Shi, et al., 2020). Following cleavage, the DNA ends are joined via non-homologous end joining (NHEJ), resulting in the precise alignment of the two coding ends and the signal ends (Rooney, et al., 2004). V(D)J recombination promotes B cell development, but aberrant V(D)J recombination can lead to precursor B-cell malignancies through RAG mediated off-target effects (Mendes, et al., 2014;Onozawa and Aplan, 2012;Thomson, et al., 2020).

The regulation of RAG expression and activity is multifactorial, serving to ensure V(D)J recombination and B cell development (Gan, et al., 2021;Kumari, et al., 2021). The RAGs consist of core and non-core region. Although non-core regions of RAG1/2 are not strictly required for V(D)J recombination, the evolutionarily conserved non-core RAG regions indicate their potential significance in vivo that may involve quantitative or qualitative modifications in RAG activity and expression (Braams, et al., 2023;Curry and Schlissel, 2008;Liu, et al., 2022;Liu, et al., 2022;Sekiguchi, et al., 2001). Specifically, the non-core RAG2 region (amino acids 384– 527 of 527 residues) contains a plant homeodomain (PHD) that can recognize histone H3K4 trimethylation, as well as a T490 locus that mediates a cell cycle-regulated protein degradation signal in proliferated pre-B cells stage (Liu, et al., 2007;

Matthews, et al., 2007). Failure to degrade RAG2 during the S stage poses a threat to the genome (Zhang, et al., 2011). Moreover, the off-target V(D)J recombination frequency is significantly higher when RAG2 is C-terminally truncated, thereby establishing a mechanistic connection between the PHD domain, H3K4me3-modified chromatin, and the suppression of off-target V(D)J recombination (Lu, et al., 2015;Mijušković, et al., 2015). The RAG1’ non-core region (amino acids1–383 of 1040 residues) has been identified as a RAG1 regulator. While the core RAG1 maintains its catalytic activity, it’s in vivo recombination efficiency and fidelity are reduced in comparison to the full-length RAG1 (fRAG1). In addition, the RAG1 binding to the genome is more indiscriminate (Beilinson, et al., 2021;Sadofsky, et al., 1993). The N-terminal domain (NTD), which is evolutionarily conserved, is predicted to contain multiple zinc-binding motifs, including a Really Interesting New Gene (RING) domain (aa 287 to 351) that can ubiquitylate various targets, including RAG1 itself (Deng, et al., 2015).. Furthermore, NTD contains a specific region (amino acids 1 to 215) that facilitates interaction with DCAF1, leading to the degradation of RAG1 in a CRL4-dependent manner (Schabla, et al., 2018). Additionally, the NTD plays a role in chromatin binding and the genomic targeting of the RAG complex (Schatz and Swanson, 2011). Despite increased evidence emphasizing the significance of non-core RAG regions, particularly RAG1’s non-core region, the function of non-core RAG regions in off-target V(D)J recombination and the underlying mechanistic basis have not been fully clarified.

Typically, genomic DNA is safeguarded against inappropriate RAG cleavage by the inaccessibility of cryptic RSSs (cRSSs), which are estimated to occur once per 600 base pairs (Teng, et al., 2015). However, recent research has demonstrated that epigenetic reprogramming in cancer can result in heritable alterations in gene expression, including the accessibility of cRSSs (Becker, et al., 2020;Fatma, et al., 2022;Goel, et al., 2022;Khoshchehreh, et al., 2019). We selected the BCR-ABL1+ B-ALL model, which is characterized by ongoing V(D)J recombinase activity and BCR-ABL1 gene rearrangement in pre-B leukemic cells (Schjerven, et al., 2017;Wong and Witte, 2004). The genome structural variations (SVs) analysis was conducted on leukemic cells from fRAG, cRAG1, and cRAG2, BCR-ABL1+ B-ALL mice to examine the involvement of non-core RAG regions in off-target V(D)J recombination events. The non-core domain deletion in both RAG1 and RAG2 led to accelerated leukemia onset and progression, as well as an increased off-target V(D)J recombination. Our analysis showed a reduction in RAG cleavage accuracy in cRAG cells and a decrease in recombinant size in cRAG1 cells, which may be responsible for the increased off-target V(D)J recombination in cRAG leukemia cells. In conclusion, our results highlight the potential importance of the non-core RAG region, particularly RAG1’s non-core region, in maintaining accuracy of V(D)J recombination and genomic stability in BCR-ABL1+ B-ALL.

Results

cRAG give more aggressive leukemia in a mouse model of BCR-ABL1+ B-ALL

In order to assess the impact of RAG activity on the clonal evolution of BCR-ABL1+ B-ALL through a genetic experiment, we utilized bone marrow transplantation (BMT) to compare disease progression in fRAG, cRAG1, and cRAG2 BCR-ABL1+ B-ALL (Schjerven, et al., 2017;Wong and Witte, 2004). Bone marrow cells transduced with a BCR-ABL1/GFP retrovirus were administered into syngeneic lethally irradiated mice, and CD19+ B cell leukemia developed within 30-80 days (Figure 1A, Supplementary Figure 1). Western blotting results confirmed equivalent transduction efficiencies of the retroviral BCR-ABL1 in all three cohorts (Supplementary Figure 2A). To investigate potential variances in leukemia outcome across different genomic backgrounds, we employed Mantel-Cox analysis to evaluate survival rates in fRAG, cRAG1, or cRAG2 mice transplanted with BCR-ABL1-transformed bone marrow cells. Our results show that, during the primary transplant phase, BCR-ABL1+ B-ALL mice expressing cRAG1 or cRAG2 demonstrated lower survival rates compared to their counterparts with fRAG (median 74.5 days versus 39 or 57 days, P < 0.0425, Figure 1A). This survival rates discrepancy was also observed during the secondary transplant phase, wherein leukemic cells were extracted from the spleens of primary recipients and subsequently purified via GFP+ cell sorting. A total of 105,104 and 103 GFP+ leukemic cells that originated from fRAG, cRAG1, or cRAG2 leukemic mice were transplanted into corresponding non-irradiated immunocompetent syngenetic recipients (survival days fRAG,11-26 days, cRAG1,10-16 days, cRAG2,11-21 days, Supplementary Figure 2B). Additionally, the cRAG mice exhibited significantly higher leukemia burdens in the peripheral blood, bone marrow, and spleen compared to the fRAG mice (Figure 1B-D). To elucidate the cellular mechanisms driving the accelerated proliferation observed in cRAG BCR-ABL1+ B-ALL, flow cytometry analyses were conducted to evaluate cell cycle dynamics and apoptotic activity. Results revealed a higher fraction of cRAG BCR-ABL1+ B-ALL cells residing in the S/G2-M phase of the cell cycle compared to their fRAG counterparts (Figure 1E). Additionally, the enhanced proliferation in cRAG leukemic cells was attributed to a reduction in apoptosis rates (Supplementary Figure 2C). RNA-seq analysis demonstrated the changes of cell differentiation and proliferation/apoptotic pathways (Supplementary Figure 3) These findings indicate that the absence of non-core RAG regions accelerates malignant transformation and leukemic proliferation, leading to a more aggressive disease phenotype in the cRAG BCR-ABL1+ B-ALL mouse model.

The loss of non-core RAG regions corresponds to a less mature cell surface phenotype but does not impede IgH VDJ recombination

To delineate the developmental stages of B cells from which the leukemic cells originated, we performed flow cytometry on single cells stained with B cell-specific surface markers. Analysis revealed that 91%-98% of GFP+ cells in cRAG mice were CD19+BP-1+B220+CD43+, indicating that most leukemic cells were at the large pre-B cell stage (Figure 2A) (Hardy and Hayakawa, 2001). Conversely, in fRAG leukemic mice, the distribution was 65% large pre-B cells (GFP+CD19+BP-1+B220+CD43+) and 35% small pre-B cells (GFP+CD19+BP-1+B220+CD43-) (Figure 2A). Moreover, approximately 5% of leukemic cells in fRAG mice expressed μHC, in con-trast to minimal expression in cRAG leukemic cells. This suggests that fRAG leukemic cells may differentiate further, associated with an immune phenotype (Figure 2B). IgH rear-rangement initiates with DH-JH joining in pro-B cells, followed by VH-DJH joining in pro-B cells, and ultimately, VL-JL rearrangements occur at the IgL loci in small pre-B cells (Schatz and Ji, 2011). Genomic PCR analysis of DNA from GFP+CD19+ cells was utilized to assess VHDJH rearrangement. The results showed a pronounced oligoclonality in cRAG leukemic cells, with tumors consistently demonstrating rear-rangements involving a restricted set of VH family members. In contrast, fRAG leukemias displayed significant polyclonality, evidenced by the widespread rear-rangement of various VH family members to all potential JH1-3 segments, indicative of a broader clonal diversity (Supplementary Figure 4). This observation aligns with the more aggressive leukemia phenotype seen in cRAG BCR-ABL1+ B-ALL mice. Such oligoclonality in cRAG leukemic cells suggests a selection process driven by BCR-ABL1-induced leukemia, favoring the emergence of a limited number of dominant leukemic clones. The absence of non-core RAG regions appears to restrict the diversity of leukemic clones, leading to the formation of oligoclonal tumors.

The loss of non-core RAG regions highlights genomic DNA damage

The findings indicate that leukemic cells from three types of mice exhibited variable arrests at the large pre-B cell stage, deviating from normal B cell developmental trajectory. Typically, at this juncture, B cells initiate the degradation of RAG2 via the cyclin-dependent kinase cyclinA/Cdk2, leading to a downregulation of RAG activity. It is therefore crucial to explore the impact of deletions in non-core regions on the expression and functionality of RAG in these leukemic cells. Analysis showed that both RAG1 (cRAG1) and RAG2 (cRAG2) were present in GFP+CD19+ splenic leukemic cells from BCR-ABL1+ B-ALL mice across different genetic backgrounds (Figure 3A). Notably, we observed an upregulation of cRAG1 and cRAG2 in leukemic cells from cRAG1 or cRAG2 mice compared to those from fRAG mice (Figure 3A, Supplementary Figure 5A). The in vitro V(D)J recombination assay confirmed that different forms of RAG exhibited cleavage activity in BCR-ABL1+B-ALL (Figure 3B and Supplementary Figure 5B).

To examine the potential correlation between aberrant RAG activity and increased DNA double-strand breaks (DSBs), we assessed levels of phosphorylated H2AX (L-H2AX), a marker of DSB response, in leukemic cells from fRAG, cRAG1, and cRAG2 mice (gated on GFP+). This evaluation aimed to gauge DNA DSBs and overall genomic instability. Flow cytometry analysis revealed elevated L-H2AX levels in cRAG1 and cRAG2 leukemic cells compared to those from fRAG, indicating a more pronounced role of cRAG in mediating somatic structural variants in BCR-ABL1+ B cells. These findings suggest enhanced expression of cRAG1 endonucleases in cRAG1 leukemic cells and increased DNA damage in cells lacking core RAG regions.

Off-target recombination mediated by RAG in BCR-ABL1+ B cells

Genome-wide sequencing and analysis were performed to compare somatic structural variants (SVs) in BCR-ABL1+ B cells derived from fRAG, cRAG1 and cRAG2 mice. The leukemic cells were sequenced with an average coverage of 25× (Supplementary Table 1). The SVs generated by RAG were screened based on two criteria: the presence of a CAC to the right (or GTG to the left) of both breakpoints, and its occurrence within 21 bp from the breakpoint (Mijušković, et al., 2015). Further elaboration on these criteria can be found in Supplementary Figure 6. Consequently, aberrant V-to-V junctions and V to intergenic regions were encompassed in five validated abnormal rearrangements at Ig loci in cRAG leukemic mice (Supplementary Table 2). Additionally, seven samples had 24 somatic structural variations, with an average of 3.4 coding region mutations per sample (range of 0-9), which is consistent with the limited number of acquired somatic mutations observed in hematological cancers (Figure 4 and Supplementary Table 3). The results of the study demonstrate that fRAG cells had low SVs (0-1 per sample), cRAG1 cells exhibited higher SVs (6-9 per sample) while cRAG2 cells had moderate SVs incidence (1-4 per sample) (Figure 4, Supplementary Table 3). These findings suggest that cRAG may lead to an elevated off-target recombination, eventually posing a threat to the BCR-ABL1+ B cells genome.

Off-target V(D)J recombination characteristics in BCR-ABL1+B cells

We further examined the characteristics of the identified structural variants (SVs). Specifically, we analyzed the exon-intron distribution profiles of 41 breakpoints from 24 SVs through genome analysis. The results indicated that 57% of the breakpoints were located within the gene body, while 43% were enriched in the flanking sequences, the majority of which were identified as transcriptional regulatory sequences (Figure 5A). P and N nucleotides are recognized as distinctive characteristics of V(D)J recombination (Repasky, et al., 2004). RSS-to-RSS and cRSS-to-cRSS recombination have P nucleotides lengths of 7 and 9, respectively, and N lengths of 5, so nucleotide lengths are basically the same during RSS-to-RSS and cRSS-to-cRSS recombination (Figure 5B). However, the frequency of P and N sequences in RSS-to-RSS recombination was 50%/50% (P/N), compared to 4%/8% (P/N) in cRSS-to-cRSS recombination (Figure 5B). This significant reduction in the frequency of P and N sequences suggests that DNA repair at off-target sites in BCR-ABL1+ B cells diverges from the classical V(D)J re-combination repair process.

The hybrid joints were notably prevalent in cRAG1 and cRAG2 leukemic cells (93% and 100%, respectively), suggesting that the non-core regions of RAG may play a role in inhibiting harmful transposition events (Figure 5C). To evaluate the effect of deleting non-core RAG regions on the emergence of oncogenic mutations, we performed a comparative analysis of cancer-related genes across three types of leukemic cells. We found that cRAG1 leukemic cells harbored a significantly higher number of cancer genes compared to the other groups. This finding corresponds with the most aggressive leukemia phenotype observed in cRAG1 BCR-ABL1+ B-ALL mice and associated changes in their transcription profiles.

The non-core regions have effects on RAG cleavage and off-target recombination size in BCR-ABL1+ B cells

Sequence logos were employed to visually contrast RSS and cRSSs within Ig and non-Ig loci, respectively. Notably, the RSS elements in Ig loci displayed a higher similarity to the canonical RSS, especially at critical functional sites. The initial four nucleotides (highlighted) of the canonical heptamer sequence CACAGTG were recognized as the cleavage site for fRAG. Conversely, in leukemic cells, the cleavage site for cRAG was pinpointed to the first three nucleotides, the CAC trinucleotide, of the heptamer sequence (Figure 6A). While both motifs (CACA and CAC) align with the highly conserved segment of the RSS heptamer sequence, differences in the cRSS sequences across off-target genes in both fRAG and cRAG mice suggest that deletion of RAG’s non-core regions broadens the spectrum of off-target substrates in BCR-ABL1+ B cells.

Antigen receptor genes are assembled by large-scale deletions and inversions (Teng, et al., 2015). The off-target recombination size was determined as the DNA fragment size spanning the two breakpoints. Our analysis demonstrated that both fRAG and cRAG2 leukemic cells produced off-target recombinations with 100% and 92% of events, respectively, spanning over 10,000 bp in length. In contrast, cRAG1 leukemic cells showed only 6% of off-target recombinations exceeding 10,000 bp, with 48% under 1,000 bp and 46% ranging between 1,000 to 10,000 bp (Figure 6BC). These findings suggest that the cRAG1 variant primarily facilitates smaller-scale off-target recombinations in BCR-ABL1+ B cells, highlighting the role of the non-core RAG1 region in influencing the extent of off-target recombination. The deletion of the non-core RAG1 region appears to constrict the size of off-target recombination, potentially contributing to the elevated frequency of off-target V(D)J recombination observed in cRAG1 leukemic cells (Figure 6D).

Discussion

In this study, we have demonstrated that non-core region deletion of both RAG1 and RAG2 leads to accelerated development of leukemia and increased off-target V(D)J recombination in mice models of BCR-ABL1+ B cells. Furthermore, we report reduced cRAG cleavage accuracy and off-target recombination size in cRAG leukemia cells, which might contribute to exacerbated off-target V(D)J recombination of cRAG BCR-ABL1+ B cells. These findings suggest that the non-core regions, particularly the non-core region of RAG1, play a crucial role in maintaining accuracy of V(D)J recombination and genomic stability in BCR-ABL1+ B cells.

Our findings suggest that leukemic cells with cRAG regions exhibit increased production of hybrid joints, implying that non-core RAG regions might suppress the formation of these hybrid joints in vivo. Post-cleavage synaptic complexes (PSCs), comprising RAG proteins, coding ends, and RSS ends, are believed to have evolved to form with optimal conformation and/or stability for conventional coding and RSS end-joining(Fugmann, et al., 2000;Libri, et al., 2021). In contrast, cRAG PSCs could promote RAG-mediated hybrid joints by facilitating closer proximity of coding and RSS ends or by increasing PSC stability. It is also conceivable that fRAGs recruit disassembly/remodeling factors to PSCs, a process that could allow non-homologous end joining (NHEJ) factors to complete the normal reaction(Fugmann, et al., 2000).

Conversely, cRAGs may have a diminished recruitment capacity due to changes in overall conformation or the absence of specific motifs, leading to more unstable PSCs and a heightened risk of accumulating incomplete hybrid joints (Raghavan, et al., 2006;Talukder, et al., 2004). Our data reveal that over 90% of junctions were hybrid joints in cRAG leukemic cells, a frequency exceeding that reported in previous studies. This suggests that deficiencies in the NHEJ pathway could contribute to chromosomal instability and lymphomagenesis(Gaymes, et al., 2002;Rassool, 2003;Scully, et al., 2019;Wiegmans, et al., 2021). Significantly, our analysis uncovered variations in the NHEJ repair pathway among leukemic cells from different genetic backgrounds, suggesting a potential aberrant expression of DNA repair pathways in cRAG leukemic cells (Supplementary Figure 3B). These findings highlight the potential of cRAG to foster increased hybrid joint formation, especially when a normal pathway for efficient coding and RSS joining is compromised in an NHEJ-aberrant context.

Our data demonstrates that the deletion of the RAG1 non-core region results in more severe off-target V(D)J recombination compared to the deletion of the RAG2 non-core region. This observation is supported by the fact that the RAG1 terminus contains multiple zinc-binding motifs and ubiquitin ligase activity, which are known to enhance the efficiency of the rearrangement reaction (Beilinson, et al., 2021;Burn, et al., 2022). Furthermore, our research reveals that RAG1 expression persists in BCR-ABL1+ progenitor B cells, and deletion of the non-core region of RAG1 results in elevated expression of RAG in comparison to fRAG. Consequently, as demonstrated in this study, cRAG1 from BCR-ABL1+ B leukemic mice is prone to generating off-target V(D)J recombination. The distinct function of RAG1’s non-core region in thymic lymphomas of Tp53-/- mice and BCR-ABL1+ B leukemic mice leads to dissimilar off-target activity of cRAG1 (Mijušković, et al., 2015). Therefore, it would be intriguing to replicate these analyses across various subtypes of ALL to further investigate this phenomenon.

In human ETV6-RUNX1 ALL, the ETV6-RUNX1 fusion gene is believed to initiate prenatally, yet the disease remains clinically latent until critical secondary events occur, leading to leukemic transformation-“pre-leukemia to leukemia” (Mijušković, et al., 2015). Genomic rearrangement, mediated by aberrant RAG recombinase activity, is a frequent driver of these secondary events in ETV6-RUNX1 ALL (Chen, et al., 2021;Papaemmanuil, et al., 2014). In contrast, RAG mediated off-target V(D)J recombination is also observed in BCR-ABL1+ B-ALL. These oncogenic structural variations can also be considered as secondary events that promote the transition - “leukemia to aggressive leukemia”. The enhancement of BCR-ABL1+ B-ALL deterioration and progression by cRAG in mouse model was consistent with our previous study that RAG enhances BCR-ABL1 positive leukemic cell growth through its endo-nuclease activity (Yuan, et al., 2021). Additionally, we showed that non-core RAG1 region deletion leads to increased cRAG1 expression and high RAG expression related to low survival in pediatric acute lymphoid leukemia (Figure 3A and Supplementary Figure 7). Therefore, more attention should be paid to the non-core RAG region mutation in BCR-ABL1+ B-ALL for the role of non-core region in leukemia suppression and off-target V(D)J recombination.

Methods

Mice

The C57BL/6 mice were purchased from the Experimental Animal Center of Xi’an Jiaotong University, while cRAG1 (amino acids 384-1040) and cRAG2 (amino acids 1-383) were obtained from Dr. David G. Schatz (Yale University, New Haven, Connecticut, USA). The mice were bred and maintained in a specific pathogen-free (SPF) environment at the Experimental Animal Center of Xi’an Jiaotong University. All animal-related procedures were in accordance with the guidelines approved by the Xi’an Jiaotong University Ethics Committee for Animal Experiments.

Generation of Retrovirus Stocks

The pMSCV-BCR-BAL1-IRES-GFP vector is capable of co-expressing the human BCR-ABL1 fusion protein and green fluorescence protein (GFP), while the pMSCV-GFP vector serves as a negative control by solely expressing GFP. To produce viral particles, 293T cells were transfected with either the MSCV-BCR-BAL1-IRES-GFP or MSCV-GFP vector, along with the packaging vector PKAT2, utilizing the X-tremeGENE HP DNA Transfection Reagent from Roche (Basel, Switzerland). After 48 hours, the viral supernatants were collected, filtered, and stored at −80°C.

Bone Marrow Transduction and Transplantation

Experiments were conducted using mice aged between 6 to 10 weeks. BCR-ABL1+ B-ALL murine model was induced by utilizing marrow from donor mice who had not undergone 5-FU treatment. The donor mice were euthanized through CO2 asphyxiation, and the bone marrow was harvested by flushing the femur and tibia with a syringe and 26-gauge needle. Erythrocytes were not removed, and 1 × 106 cells per well were plated in six-well plates. A single round of co-sedimentation with retroviral stock was performed in medium containing 5% WEHI-3B-conditioned medium and 10 ng/mL IL-7 (Peprotech, USA). After transduction, cells were either transplanted into syngeneic female recipient mice (1 × 106 cells each) that had been lethally irradiated (2 × 450 cGy), or cultured in RPMI-1640 (Hyclone, Logan, UT) medium supplemented with 10% fetal calf serum (Hyclone), 200 mmol/L L-glutamine, 50 mmol/L 2-mercaptoethanol (Sigma, St Louis, MO), and 1.0 mg/mL penicillin/streptomycin (Hyclone). Subsequently, recipient mice were monitored daily for indications of morbidity, weight loss, failure to thrive, and splenomegaly. Weekly assessment of peripheral blood GFP percentage was done using FACS analysis of tail vein blood. Hematopoietic tissues and cells were utilized for histopathology, in vitro culture, FACS analysis, secondary transplantation, genomic DNA preparation, protein lysate preparation, or lineage analysis, contingent upon the unique characteristics of mice under study.

Secondary Transplants

Thawed BM cells were sorted using a BD FACS Aria II (Becton Dickinson, San Jose California, USA). GFP positive leukemic cells (1× 106, 1×105, 1×104, and 1×103) were then resuspended in 0.4 mL Hank’s Balanced Salt Solution (HBSS) and intravenously administered to unirradiated syngeneic mice.

Flow cytometry analysis and sorting

Bone marrow, spleen cells, and peripheral blood were harvested from leukemic mice. Red blood cells were eliminated using NH4Cl RBC lysis buffer, and the remaining nucleated cells were washed with cold PBS. In order to conduct in vitro cell surface receptor staining, 1× 106 cells were subjected to antibody staining for 20 minutes at 4°C in 1×PBS containing 3% BSA. Cells were then washed with 1×PBS and analyzed using a CytoFLEX Flow Cytometer (Beckman Coulter, Miami, FL) or sorted on a BD FACS Aria II. Apoptosis was analysed by resuspending the cells in Binding Buffer (BD Biosciences, Baltimore, MD, USA), and subsequent labeling with antiannexin V-AF647 antibody (BD Biosciences) and propidium iodide (BD Biosciences) for 15 minutes at room temperature. The lineage analysis was performed using the following antibodies, which were purchased from BD Biosciences: anti-BP-1-PerCP-Cy7, anti-CD19-PerCP-CyTM5.5, anti-CD43-PE, anti-B220-APC, and anti-μHC-APC.

BrdU incorporation and analysis

Cells obtained from primary leukemic mice were cultured in six-well plates containing RPMI-1640 medium supplemented with 10% FBS and 50 µg/mL BrdU. After incubation at 37°C for 30 minutes, the cells were harvested and intranuclearly stained with anti-BrdU and 7-AAD antibodies, following the manufacturer’s instructions.

The in vitro V(D)J recombination assay

The retroviral recombination substrate pINV-12/23 was introduced into primary leukemic cells utilizing X-treme GENE HP DNA Transfection Reagent (Roche). Recombina-tion efficiency of pINV-12/23 was evaluated through flow cytometry analysis for mouse CD90 (mCD90) and hCD4 expression (Yuan, et al., 2021).

Western blotting analysis

Over 1 × 106 leukemic cells were centrifuged and washed with ice-cold PBS. The cells were then treated with ice-cold RIPA buffer, consisting of 50 mM Tris-HCl (pH 7.4), 0.15 M NaCl, 1% Triton X-100, 0.5% NaDoc, 0.1% sodium dodecyl sulphide (SDS), 1 mM ethylene diamine tetraacetic acid (EDTA), 1 mM phenylmethane sulphony fluoride (PMSF) (Amresco), and fresh protease inhibitor cocktail Pepstain A (Sigma). After sonication using a Bioruptor TMUCD-200 (Diagenode, Seraing, Belgium), the suspension was spined at 14,000 g for 3 minutes at 4°C. The total cell lysate was either utilized immediately or stored at −80°C. Protein concentrations were determined using DC Protein Assay (Bio–Rad Laboratories, Hercules, California, USA). Subsequently, the protein samples (20 μg) were incubated with α-RAG1 (mAb 23) and α-RAG2 (mAb 39) antibodies (Teng, et al., 2015), with GAPDH serving as the loading control. The signal was further detected using secondary antibody of goat anti-rabbit IgG conjugated with horseradish peroxidase (Thermo Scientific, Waltham, MA). The band signal was developed with Immobilon™ Western Chemiluminescent HRP substrate (Millipore, Billerica, MA). The band development was analyzed using GEL-PRO ANALYZER software (Media Cybernetics, Bethesda, MD).

Genomic PCR

Genomic PCR was performed in a 20μl reaction containing 50 ng of genomic DNA, 0.2 μm of forward and reverse primer, and 10 μl Premix Ex Taq (TaKaRa, Shiga, Japan). Amplification conditions were as follows: 94°C for 5 minutes; 35 cycles of 30 seconds at 94°C, 30 seconds at 60°C and 1 minutes at 72°C; 72°C for 5 minutes (BioRad, Hercules, CA). Genomic PCR was performed using the following primers: DHL-5′-GGAATTCGMTTTTTGTSAAGGGATCTACTACTGTG-3′;JH3-5′-GTCTAGATTCTCAC-AAGAGTCCGATAGACCCTGG-3′; VHQ52-5′-CGGTACCAGACTGARCATCASCAAG-GACAAYTCC-3′; VH558-5′-CGAGCTCTCCARCACAGCCTWCATGCARCTCARC-3′; VH7183-5′-CGGTACCAAGAASAMCCTGTWCCTGCAAATGASC-3′.(Schlissel, et al., 1991):

RNA-seq library preparation and sequencing

GFP+CD19+ cells were sorted from the spleen of cRAG1 (n=3, 1×106 cells /sample), cRAG2 (n=3, 1×106 cells /sample), and fRAG (n=3, 1×106 cells /sample) BCR-ABL1+ B-ALL mice. Total RNA was extracted using Trizol reagent (Invitrogen, CA, USA) following the manufacturer’s guidelines. RNA quantity and purity analysis was done using Bioanalyzer 2100 and RNA 6000 Nano LabChip Kit (Agilent, CA, USA) with RIN number >7.0. RNA-seq libraries were prepared by using 200 ng total RNA with TruSeq RNA sample prep kit (Illumina). Oligo(dT)-enriched mRNAs were fragmented randomly with fragmentation buffer, followed by first- and second-strand cDNA synthesis. After a series of terminal repair, the double-stranded cDNA library was obtained through PCR enrichment and size selection. cDNA libraries were sequenced with the Illumina Hiseq 2000 sequencer (Illumina HiSeq 2000 v4 Single-Read 50 bp) after pooling according to its expected data volume and effective concentration. Two biological replicates were performed in the RNA-seq analysis. Raw reads were then aligned to the mouse genome (GRCm38) using Tophat2 RNA-seq alignment software, and unique reads were retained to quantify gene expression counts from Tophat2 alignment files. The differentially expressed mRNAs and genes were selected with log2 (fold change) >1 or log2 (fold change) <-1 and with statistical significance (p value < 0.05) by R package. Bioinformatic analysis was performed using the OmicStudio tools athttps://www.omicstudio.cn/tool.

Preparation of tumor DNA samples

GFP+CD19+ splenic cells, tail and kidney tissue were obtained from cRAG1, cRAG2 and fRAG BCR-ABL1+ B-ALL mice, and genomic DNA was extracted using a TIANamp Genomic DNA Kit (TIANGEN-DP304). Subsequently, paired-end libraries were constructed from 1 µg of the initial genomic material using the TruSeq DNA v2 Sample Prep Kit (Illumina, #FC-121-2001) as per the manufacturer’s instructions. The size distribution of the libraries was assessed using an Agilent 2100 Bioanalyzer (Agilent Technologies, #5067-4626), and the DNA concentration was quantified using a Qubit dsDNA HS Assay Kit (Life Technologies, #Q32851). The Illumina HiSeq 4000 was utilized to sequence the samples, with two to four lanes allocated for sequencing the tumor and one lane for the control DNA library of the kidney or liver, each with 150 bp paired end reads.

Read alignment and structural variant calling

Fastq files were generated using Casava 1.8 (Illumina), and BWA 37 was employed to align the reads to mm9. PCR duplicates were eliminated using Picard’s Mark Duplicates tool (source-forge.net/apps/mediawiki/picard). Our custom scripts (http://sourceforge.net/projects/svdetection) were utilized to eliminate BWAdesignated concordant and read pairs with low BWA mapping quality scores. Intrachromosomal and inter-chromosomal rearrangements were identified using SV Detect from discordant, quality prefiltered read pairs. The mean insertion size and standard deviation for this analysis were obtained through Picard’s InsertSizeMetrics tool (sourceforge.net/apps/mediawiki/picard). Tumor-specific structural variants (SVs) were identified using the manta software (https://github.com/Illumina/manta/blob/mater/docs/userGuide/README.md#introduction).

Validation of high confidence off-target candidates

The elimination of non-specific structural mutations from the kidney or tail was necessary for tumor-specific structural variants identification. Subsequently, the method involving 21-bp CAC-to-breakpoint was employed to filter RAG-mediated off-target gene. The validation of high confidence off-target candidates was carried out through PCR. Oligonucleotide primers were designed to hybridize within the “linking” regions of SV Detect, in the appropriate orientation. The PCR product was subjected to Sanger sequencing and aligned to the mouse mm9 reference genome using BLAST (https://blast.ncbi.nlm.nih.gov/Blast.cgi).

Statistics

Statistical analysis was conducted using SPSS 20.0 (IBM Corp.) and GraphPad Prism 6.0 (GraphPad Software). Descriptive statistics were reported as means ± standard deviation for continuous variables. Statistical analyses were applied to biologically independent mice or technical replicates for each experiment which was independently repeated at least three times. The equality of variances was assessed using Levene’s test. Two-group comparisons, multiple group comparisons, and survival comparisons were performed using independent-samples t-test, one-way analyses of variance (ANOVA) with post hoc Fisher’s LSD test, and log-rank Mantel-Cox analysis, respectively. Kaplan-Meier survival curves were utilized to depict the changes in survival rate over time. Statistical significance was set at P<0.05.

Supporting information